---
title: "FragPipeAnalystR DIA Global Proteome Tutorial"
output:
  html_document:
    keep_md: yes
---

# Introduction

FragPipeAnalystR is a R package intended for downstream analysis of data generated from [FragPipe](https://fragpipe.nesvilab.org/). Here we demonstrated the utility of FragPipeAnalyst by reanalyzing a clear cell renal cell carcinoma (ccRCC) data-independent acquisition (DIA) data collected by CPTAC. You can download the example files from [here](https://zenodo.org/records/13147961). Files are in "DIA_4plex" folder.

## CCRCC DIA data

As described in the manuscript, DIA ccRCC data were fetched from [Clark et al. (2019)](https://doi.org/10.1016/j.cell.2019.10.007) and processed via [FragPipe](https://fragpipe.nesvilab.org/). As you will see in the following sections. The result is quite similar to corresponding TMT data.

``` {r include=T, warning=F}
library(FragPipeAnalystR)
ccrcc <- make_se_from_files("/Users/hsiaoyi/Documents/workspace/FragPipeR_manuscript/data/DIA_4plex/diann-output.pg_matrix.tsv",
                         "/Users/hsiaoyi/Documents/workspace/FragPipeR_manuscript/data/DIA_4plex/experiment_annotation_clean.tsv",
                         type = "DIA")
```

``` {r include=T, warning=F}
print(head(rownames(ccrcc)))
```

``` {r include=T, warning=F}
plot_pca(ccrcc)
```

``` {r include=T, warning=F}
plot_correlation_heatmap(ccrcc)
```

``` {r include=T, warning=F}
plot_missval_heatmap(ccrcc)
```

``` {r include=T, warning=F}
plot_feature_numbers(ccrcc)
```

```{r include=T, warning=F}
plot_feature(ccrcc, c("Q16790", # CA9
                      "Q8IVF2", # AHNAK2
                      "P19404", # NDUFV2
                      "P01833" # PIGR
                      ))
```

## Imputation

``` {r include=T, warning=F}
imputed <- manual_impute(ccrcc)
```

``` {r include=T, warning=F}
plot_pca(imputed)
```

```{r include=T, warning=F}
plot_feature(imputed,  c("Q16790", # CA9
                      "Q8IVF2", # AHNAK2
                      "P19404", # NDUFV2
                      "P01833" # PIGR
                      ))
```

## DE result without imputation
```{r include=T, warning=F}
de_result <- test_limma(ccrcc, type = "all")
de_result_updated <- add_rejections(de_result)
plot_volcano(de_result_updated, "Tumor_vs_NAT")
```

## DE result with imputation
```{r include=T, warning=F}
de_result <- test_limma(imputed, type = "all")
de_result_updated <- add_rejections(de_result)
plot_volcano(de_result_updated, "Tumor_vs_NAT")
```

One of the differences between differential expression analysis results from unimputed and imputed data is CA9 (Carbonic anhydrase 9, Q16790) which is a [known marker of clear cell renal cell carcinoma](https://doi.org/10.1016/j.ejca.2010.07.020). Its expression is missing in the normal data, which leads its absence of unimputed data.

``` {r include=T, warning=F}
sessionInfo()
```
