---
title: "FragPipeAnalystR LFQ tutorial"
output:
  html_document:
    keep_md: yes
---

# Introduction

Here we presented the usage of FragPipeAnalystR to perform AP-MS analysis in an alternative way. Previously, the protein-level analysis was demonstrated in the manuscript via FragPipe-Analyst website, but here we are going to analyze it at the peptide-level.

## Reading input files
``` {r include=T, warning=F}
library(FragPipeAnalystR)
se <- make_se_from_files("/Users/hsiaoyi/Documents/workspace/FragPipeR_manuscript/data/AP-MS/combined_peptide.tsv",
                         "/Users/hsiaoyi/Documents/workspace/FragPipeR_manuscript/data/AP-MS/experiment_annotation.tsv",
                         type = "LFQ", level = "peptide")
```

``` {r include=T, warning=F}
print(head(rownames(se)))
``` 

``` {r include=T, warning=F}
plot_pca(se)
```

``` {r include=T, warning=F}
plot_correlation_heatmap(se)
```

``` {r include=T, warning=F}
plot_missval_heatmap(se)
```

``` {r include=T, warning=F}
plot_feature_numbers(se)
```

```{r include=T, warning=F}
colData(se)$condition
```

## Imputation
```{r include=T, warning=F}
imputed_se <- manual_impute(se)
```


``` {r include=T, warning=F}
plot_pca(imputed_se)
```

``` {r include=T, warning=F}
plot_correlation_heatmap(imputed_se)
```

## Differential expression analysis
```{r include=T, warning=F}
de_result <- test_limma(imputed_se, type = "all")
de_result_updated <- add_rejections(de_result)
```

``` {r include=T, warning=F}
plot_volcano(de_result_updated, "CCND1_vs_CONTROL")
```

The volcano could be labelled in a different way via `name_col` argument of the function:

``` {r include=T, warning=F}
plot_volcano(de_result_updated, "CCND1_vs_CONTROL", name_col="Gene")
```

``` {r include=T, warning=F}
sessionInfo()
```
